---
title: "Assesment1"
author: "Volodymyr Dovhanyk"
date: "Friday, August 14, 2015"
output: html_document
---
#Loading and preprocessing the data
This markupfile is generated as resualts of peer assesment for week 2 at Coursera "Reproducible research" courses.
Firstly we download th file with initial data into your working directory and read the data with R. 
**This code can produce some warnings but file should besuccesfully downloaded and precessed.**

```{r results='hide', cache=TRUE}
unzip (zipfile="activity.zip", exdir="data")
data<-read.csv("data/activity.csv")
```
#What is mean total number of steps taken per day?
The mean of steps per day are presented below
```{r meansteps}
meanSteps<-aggregate(data$steps, by=list(data$date), FUN=mean)
```
In addition we can evaluate the mean values based on the histogram.
*Please note NAs are replaced by 0*
```{r}
meanSteps$x[is.na(meanSteps[2])]<-0
hist(as.numeric(meanSteps$x), xlab="Mean steps per day", main="Mean steps per day")
```
``` {r meanval, echo=FALSE}
meanval<-mean(data$steps, na.rm=T)
```
```{r medianval, echo=FALSE}
medianval<-median(data$steps)
```
The mean  of the total number of steps taken per day is equal `r meanval` and median `r medianval`. 

#What is the average daily activity pattern?
Average number of steps taken, averaged across all days is presented in the chart below

```{r}
meanStepsDayly<-aggregate(data$steps, by=list(data$interval), FUN=mean, na.rm=T)
plot(meanStepsDayly$Group.1, meanStepsDayly$x, type="l", 
     xlab="Interval", ylab="Step mean")
```
```{r bestInterval, echo=FALSE,results='hide'}
best<-meanStepsDayly[meanStepsDayly[2]==max(meanStepsDayly[2])]
bestInterval<-round(best[1],0)
bestSteps<-best[2]
```
Maximum number of steps os reached at `r bestInterval` and reaches value `r bestSteps`
#Imputing missing values
``` {r echo=FALSE, results='hide'}
na.num<-sum(is.na(data$steps))

```
Initial dataset contains `r na.num` NA values.
To improve quality of dataset we create new dataset CleanData and replace NAs with mean values for the corresponding intervals.

``` {r results='hide'}
CleanData<-data
for (i in 1:length(CleanData$steps)) {
    if (is.na(CleanData$steps[i])) {
        CleanData$steps[i]<-
            meanStepsDayly$x[meanStepsDayly$Group.1==CleanData$interval[i]]
    }
}
```

Results of such kind repacement can be evaluated on the cart bellow.
``` {r}
NewMeanSteps<-aggregate(CleanData$steps, by=list(data$date), FUN=mean)
hist(as.numeric(NewMeanSteps$x), xlab="Mean steps per day", main="Mean steps per day (cleaned data)")

NewMean<-mean(CleanData$steps)
NewMedian<-median(CleanData$steps)
```

After replacement the steps mean was increased to `r NewMean` and median value is the same `r NewMedian` because of huge numbers of zeros in initial dataset.

#Are there differences in activity patterns between weekdays and weekends?
First of all we defined weekdays and append corresponding column to the cleaned dataset.
```{r results='hide'}
weekdays<-factor(c("weekend","weekday"))
CleanData<-cbind(CleanData,weekdays)
weekend<-c("субота", "нед≥л€")
weekday<-c("понед≥лок", "в≥второк", "середа", "четвер", "п'€тниц€")
CleanData$weekdays[weekdays(as.Date(CleanData$date)) %in% weekday]<-"weekday"
CleanData$weekdays[weekdays(as.Date(CleanData$date)) %in% weekend]<-"weekend"

```
Please notice for weekdays we use Ukranian names `r weekend` and `r weekday` instead of English ones, because of using cyrylic settings in OS.
We can evaluate the differences between weekdyas and weekends in the chart below.
```{r}
meanStepsDaylyWD<-aggregate(CleanData$steps, 
                            by=list(CleanData$interval, CleanData$weekdays), FUN=mean)

par(mfrow=c(2,1))

plot(x=meanStepsDaylyWD$Group.1[meanStepsDaylyWD$Group.2=="weekday"],
     y=meanStepsDaylyWD$x[meanStepsDaylyWD$Group.2=="weekday"], type="l", 
     xlab="intervals", ylab="Mean steps", main="Weekdays")
plot(x=meanStepsDaylyWD$Group.1[meanStepsDaylyWD$Group.2=="weekend"],
     y=meanStepsDaylyWD$x[meanStepsDaylyWD$Group.2=="weekend"], type="l", 
     xlab="intervals", ylab="Mean steps", main="Weekends")
```

